---
title: "Cleaning_Data"
output: html_document
date: "2026-03-28"
---

## Script Description
This script creates daily data for the load, relative humidity, and temperature datasets. The 'data'_raw are the raw datasets, while the 'name'_proc are the processed daily data. 

The processed datasets 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results = 'hide', message = FALSE}

## loading libraries

library(forecast)
library(lubridate)
library(ggplot2)
library(forecast)
library(Kendall)
library(tseries)
library(tidyverse)
library(readxl)
library(dplyr)


```

```{r}
#Inspecting Raw Data

load_raw <- read_xlsx("./Data/Raw/load.xlsx")
humid_raw <- read_xlsx("./Data/Raw/relative_humidity.xlsx")
temp_raw <- read_xlsx("./Data/Raw/temperature.xlsx")

head(load_raw)
head(humid_raw)
head(temp_raw)


```
Load data has hourly observations organized by columns, while the temperature and humidity data have hourly data in columns. For load data will merge across columns, but for the temp + humidity will group by date and average the group values by column. 


```{r}
## Cleaning Load Data

load_proc <- load_raw %>% 
  mutate(daily_ave = rowMeans(load_raw[ ,c(3:26)], na.rm = TRUE)) %>% 
  select(date, daily_ave)


head(load_proc)
tail(load_proc)



```
Dataset is cleaned between Jan 2005 - June 2011
For the project, we need to forecast january and feb of 2011. Will need a ts object with jan 2005-dec2010

```{r}
# Cleaning Humidity Data

humid_proc <- humid_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))
    

tail(humid_proc)


```


```{r}

# Cleaning Temperature Data

temp_proc <- temp_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))

#removing NA row at the end of the data
temp_rows <- nrow(temp_proc) - 1
temp_proc <- temp_proc[1:temp_rows, ]

temp_proc
tail(temp_proc)

```


```{r}
#Checking all the rows are the same length
#after checking head/ tail of individual datasets -- to check if start & end dates line up

nrow(temp_proc)
nrow(load_proc)
nrow(humid_proc)

```



```{r}
write_csv(temp_proc, "./Data/Daily_Data/Temperature_Daily.csv")
write_csv(humid_proc, "./Data/Daily_Data/Humid_Daily.csv")
write_csv(load_proc, "./Data/Daily_Data/Load_Daily.csv")


```






