---
title: "Cleaning_Data"
output: html_document
date: "2026-03-28"
---

## Script Description
This script creates daily data for the load, relative humidity, and temperature datasets. The 'data'_raw are the raw datasets, while the 'name'_proc are the processed daily data. 


## From Luana's Instructions: 
For model training, use data from January 1, 2005, to June 30, 2010. Once you have processed the data, fit models using this training period. To evaluate model performance, use the fitted models to forecast daily demand for July 1, 2010, to July 31, 2010, as a validation set. Assess which model(s) provide accurate forecasts.

After selecting the best-performing model(s), re-train the model using the full dataset from January 1, 2005, to June 30, 2011. Use this updated model to generate forecasts for July 1, 2011, to July 31, 2011. The final forecasts will be submitted as part of the competition.

## 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results = 'hide', message = FALSE}

## loading libraries

library(forecast)
library(lubridate)
library(ggplot2)
library(forecast)
library(Kendall)
library(tseries)
library(tidyverse)
library(readxl)
library(dplyr)


```

```{r}
#Inspecting Raw Data

load_raw <- read_xlsx("./Data/Raw/load.xlsx")
humid_raw <- read_xlsx("./Data/Raw/relative_humidity.xlsx")
temp_raw <- read_xlsx("./Data/Raw/temperature.xlsx")

head(load_raw)
head(humid_raw)
head(temp_raw)


```
Load data has hourly observations organized by columns, while the temperature and humidity data have hourly data in columns. For load data will merge across columns, but for the temp + humidity will group by date and average the group values by column. 


```{r}
## Cleaning Load Data

load_proc <- load_raw %>% 
  mutate(daily_ave = rowMeans(load_raw[ ,c(3:26)], na.rm = TRUE)) %>% 
  select(date, daily_ave)


head(load_proc)
tail(load_proc)



```
Dataset is cleaned between Jan 2005 - June 2011
For the project, we need to forecast january and feb of 2011. Will need a ts object with jan 2005-dec2010

```{r}
# Cleaning Humidity Data

humid_proc <- humid_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))
    

tail(humid_proc)


```


```{r}

# Cleaning Temperature Data

temp_proc <- temp_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))

#removing NA row at the end of the data
temp_rows <- nrow(temp_proc) - 1
temp_proc <- temp_proc[1:temp_rows, ]

temp_proc
tail(temp_proc)

```


```{r}
#Checking all the rows are the same length
#after checking head/ tail of individual datasets -- to check if start & end dates line up

nrow(temp_proc)
nrow(load_proc)
nrow(humid_proc)

```


Un-comment the following out to re-save the csv files
```{r}
#write_csv(temp_proc, "./Data/Daily_Data/Temperature_Daily.csv")
#write_csv(humid_proc, "./Data/Daily_Data/Humid_Daily.csv")
#write_csv(load_proc, "./Data/Daily_Data/Load_Daily.csv")


```


Creating the model training data subset (Jan 2005 through July 2010)
```{r}
temp_train_rows <- nrow(temp_proc) - 334
temp_training <- temp_proc[1: temp_train_rows, ]

humid_train_rows <- nrow(humid_proc) - 334
humid_training <- humid_proc[1: humid_train_rows, ]

load_train_rows <- nrow(load_proc) - 334
load_training <- load_proc[1: load_train_rows, ]

tail(temp_training)
tail(humid_training)
tail(load_training)

```


```{r}
write_csv(temp_training, "./Data/Model_Training/Temperature_Train.csv")
write_csv(humid_training, "./Data/Model_Training/Humid_Train.csv")
write_csv(load_training, "./Data/Model_Training/Load_Train.csv")

```





