---
title: "NG_Forecast"
output:
  pdf_document: default
  html_document: default
date: "2026-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(forecast)
library(Kendall)
library(tseries)
library(tidyverse)
library(readxl)
library(dplyr)
library(cowplot)
library(lubridate)
library(ggplot2)
library(smooth)
library(zoo)
library(kableExtra)

```

Initial Steps
1. plots
2. acf/ Pacf
3. trend
4. Seasonality
5. Stationary tests
6. Outliers
7. Missing data



```{r}
#full data until July 31
load_data <- read_csv("./Data/Model_Training/Load_Train.csv")


#subset of data until June 30 (used for training the mode) - July 1-31 we will be forecasting
train_rows <- nrow(load_data) - 31


load_subset <- load_data[1:train_rows, ]

#checking the data ends on the right date
tail(load_data)
tail(load_subset)
head(load_subset)

```

```{r}
temp_data <- read_csv("./Data/Model_Training/Temperature_Train.csv")

humid_data <- read_csv("./Data/Model_Training/Humid_Train.csv")

head(temp_data)
head(humid_data)

head(load_data)


```




# Creating Time Series Objects - Multiple Seasonality

```{r}
full_load_ts <- msts(load_data$daily_ave, 
                     seasonal.periods = c(7, 365.25), 
                      start=c(2005,01,01))

# creating a testing subset
n_for <- 31
full_len <- nrow(load_data)
train_data <- load_data[1:(full_len - n_for), ]
test_data  <- load_data[(full_len - n_for + 1):full_len, ]

train_load_ts <- msts(train_data$daily_ave,
                      seasonal.periods = c(7, 365.25),
                      start = c(2005, 1, 1))

test_load_ts <- msts(test_data$daily_ave,
                    seasonal.periods = c(7, 365.25),
                    start = end(train_load_ts) + c(0,1))
```





```{r}
#Checking to make sure that the testing/ training subsets were lined up in the correct locations. Full ts was plotted underneath layers. 

autoplot(full_load_ts, series = "full ts")+
  autolayer(train_load_ts, series = "training ts")+
  autolayer(test_load_ts, series = "test")

```



## STL + ETS Model

```{r}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(train_load_ts,h=31)

#Plot foresting results
autoplot(ETS_fit) + ylab("Load")

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Load")
```
 
## ARIMA + Fourier terms model

```{r}

ARIMA_Four_fit <- auto.arima(train_load_ts, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(train_load_ts, 
                                          K=c(2,12))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(train_load_ts,
                                        K=c(2,12),
                                        h=31),
                           h=31
                           ) 

#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Load")

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Load")


```

 
## TBATS Model

```{r}

# TBATS can take time to fit
TBATS_fit <- tbats(train_load_ts)

TBATS_for <- forecast(TBATS_fit, h=31)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Load") 

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)+
  ylab("Load") 


```



## Neural Network Fit
```{r}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(train_load_ts,
                 p=1,
                 P=0,
                 xreg=fourier(train_load_ts, K=c(2,12)))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_fit, 
                   h=31,
                   xreg=fourier(train_load_ts, 
                                          K=c(2,12),h=31))


#Plot foresting results
autoplot(NN_for) +
  ylab("Load") 


#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Load") 
```



## Incorporating Weather data

```{r}
temp_pc1 <- prcomp(temp_data %>% select(-date), scale. = TRUE)$x[,1]
hum_pc1  <- prcomp(humid_data %>% select(-date), scale. = TRUE)$x[,1]

weather_df <- data.frame(
  date = temp_data$date,
  temp = temp_pc1,
  humidity = hum_pc1
)

train_data <- left_join(train_data, weather_df, by = "date")
test_data  <- left_join(test_data, weather_df, by = "date")


xreg_train <- as.matrix(train_data %>% select(temp, humidity))
xreg_test  <- as.matrix(test_data %>% select(temp, humidity))

sarimax_fit <- auto.arima(
  train_data$daily_ave,
  xreg = xreg_train,
  seasonal = TRUE,
  stepwise = FALSE,
  approximation = FALSE
)

sarimax_for <- forecast(
  sarimax_fit,
  xreg = xreg_test,
  h = 31
)

sarimax_ts <- ts(
  sarimax_for$mean,
  start = end(train_load_ts) + c(0, 1),
  frequency = frequency(train_load_ts)
)

```


# Just temperature SARIMAX
```{r}
sarimax_temp_fit <- auto.arima(
  train_data$daily_ave,
  xreg = train_data$temp,
  seasonal = TRUE,
  stepwise = FALSE,
  approximation = FALSE
)

sarimax_temp_for <- forecast(
  sarimax_temp_fit,
  xreg = test_data$temp,
  h = 31
)

sarimax_temp_ts <- ts(
  sarimax_temp_for$mean,
  start = end(train_load_ts) + c(0,1),
  frequency = frequency(train_load_ts)
)



```



#just humidity sarimax
```{r}

sarimax_hum_fit <- auto.arima(
  train_data$daily_ave,
  xreg = train_data$humidity,
  seasonal = TRUE,
  stepwise = FALSE,
  approximation = FALSE
)

sarimax_hum_for <- forecast(
  sarimax_hum_fit,
  xreg = test_data$humidity,
  h = 31
)

sarimax_hum_ts <- ts(
  sarimax_hum_for$mean,
  start = end(train_load_ts) + c(0,1),
  frequency = frequency(train_load_ts)
)



```



## Combining TBATS and NN

```{r}
#averaging of functions
ensemble_mean <- 0.9 * TBATS_for$mean + 0.1 * NN_for$mean

ensemble_ts <- ts(
  ensemble_mean,
  start = start(test_load_ts),
  frequency = frequency(test_load_ts)
)
```





## comparing forecasts
```{r}
#Model 1: STL + ETS
ETS_scores <- accuracy(ETS_fit$mean,test_load_ts)  

#Model 2: ARIMA + Fourier 
ARIMA_scores <- accuracy(ARIMA_Four_for$mean,test_load_ts)

# Model 3:  TBATS 
TBATS_scores <- accuracy(TBATS_for$mean,test_load_ts)

# Model 4:  Neural Network 
NN_scores <- accuracy(NN_for$mean,test_load_ts)

#SARIMAX 1
sarimax_scores <- accuracy(sarimax_ts,test_load_ts)

#averaging test
tbats_nn_mix_score <- accuracy(ensemble_ts, test_load_ts)

#Just temp sarimax
temp_sarimax <- accuracy(sarimax_temp_ts, test_load_ts)

#just humidity sarimax
humid_sarimax <- accuracy(sarimax_hum_ts, test_load_ts)


```


#comparing performance metrics

```{r}
#create data frame
scores <- as.data.frame(
  rbind(ETS_scores, ARIMA_scores, TBATS_scores, NN_scores, sarimax_scores, tbats_nn_mix_score, temp_sarimax, humid_sarimax)
  )
row.names(scores) <- c("STL+ETS", "ARIMA+Fourier","TBATS","NN", "Sarimax", "TBATS+NN", "Temp SARIMAX", "Humid Sarimax")

#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))     
```


## creating a table of scores
```{r}
kbl(scores, 
      caption = "Forecast Accuracy for Load",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(scores[,"RMSE"]))
```


#plotting the different functions
```{r}
autoplot(train_load_ts) +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for,PI=FALSE, series="TBATS") +
  autolayer(NN_for,PI=FALSE, series="NN") +
  autolayer(sarimax_ts, PI = FALSE, series = "SARIMAX")+
  xlab("Day") + ylab("Daily Active Power") +
  guides(colour=guide_legend(title="Forecast"))
```

## taking a closer look at the forecasts
```{r}
autoplot(test_load_ts) +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for,PI=FALSE, series="TBATS") +
  autolayer(NN_for,PI=FALSE, series="NN")+
  autolayer(sarimax_ts, PI = FALSE, series = "SARIMAX")+
  ylab("Daily Active Power") +
  guides(colour=guide_legend(title="Forecast"))
``` 
Out of the four multi-seasonality forecast models, the TBATS model had the best performance across metrics except for MPE. For the first forecast submission, the TBATS model will be selected. 


### Saving of forecasts

 

```{r}
template <- read_csv("submission_template.csv")
head(template)
names(template)
```


```{r}
submission <- template %>%
  mutate(load = as.numeric(TBATS_for$mean))

write_csv(submission, "./Forecast_Submissions/TBATS_submission.csv")

submission <- template %>%
  mutate(load = as.numeric(NN_for$mean))

write_csv(submission, "./Forecast_Submissions/NN_submission.csv")

submission <- template %>%
  mutate(load = as.numeric(ARIMA_Four_for$mean))

write_csv(submission, "./Forecast_Submissions/ARIMA_Four_for_submission.csv")

submission <- template %>%
  mutate(load = as.numeric(sarimax_ts))

write_csv(submission, "./Forecast_Submissions/Sarimax_for_submission.csv")

submission <- template %>%
  mutate(load = as.numeric(ETS_fit$mean))

write_csv(submission, "./Forecast_Submissions/ETS_for_submission.csv")



```


