
---
title: "Gutkowski Shen Forecast Submission"
output: 
  html_document: default
  pdf_document: default
date: "`r Sys.Date()`"
---

# 1. Introduction

The objective of this project is to forecast daily electrical load using historical load data and meteorological variables. Multiple time series models were evaluated, including seasonal decomposition methods, ARIMA-based models, neural networks, and models incorporating exogenous weather variables. Model performance was assessed using RMSE and Kaggle competition scores.


# 2. Data Import and Wrangling

```{r setup, message=FALSE, warning=FALSE}
library(forecast)
library(lubridate)
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)
library(kableExtra)

### Load and process raw data

load_raw <- read_xlsx("./Data/Raw/load.xlsx")
humid_raw <- read_xlsx("./Data/Raw/relative_humidity.xlsx")
temp_raw <- read_xlsx("./Data/Raw/temperature.xlsx")

# Process load data
load_proc <- load_raw %>% 
  mutate(daily_ave = rowMeans(load_raw[,3:26], na.rm = TRUE)) %>% 
  select(date, daily_ave)

# Process humidity
humid_proc <- humid_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))

# Process temperature
temp_proc <- temp_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))

# Create time series
full_load_ts <- msts(load_proc$daily_ave, 
                     seasonal.periods = c(7, 365.25), 
                     start=c(2005,1,1))

# Train/test split
n_for <- 31
full_len <- nrow(load_proc)

train_data <- load_proc[1:(full_len - n_for), ]
test_data  <- load_proc[(full_len - n_for + 1):full_len, ]

train_load_ts <- msts(train_data$daily_ave,
                      seasonal.periods = c(7, 365.25),
                      start = c(2005,1,1))

test_load_ts <- msts(test_data$daily_ave,
                     seasonal.periods = c(7, 365.25),
                     start = end(train_load_ts) + c(0,1))
```

# 3. Models

## 3.1 TBATS Model
```{r}

TBATS_fit <- tbats(train_load_ts)
TBATS_for <- forecast(TBATS_fit, h=31)

autoplot(TBATS_for) + ylab("Load")

```

## 3.2  TBATS + NN Combination (50/50 weighting)
```{r}
#averaging of functions
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(train_load_ts,
                 p=1,
                 P=0,
                 xreg=fourier(train_load_ts, K=c(2,12)))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_fit, 
                   h=31,
                   xreg=fourier(train_load_ts, 
                                          K=c(2,12),h=31))



combo3_mean <- 0.5 * TBATS_for$mean + 0.5 * NN_for$mean

combo3_mean <- ts(
  combo3_mean,
  start = start(test_load_ts),
  frequency = frequency(test_load_ts)
)

```

## 3.3 TBATS + Neural Network Combination

```{r}
combo_mean <- 0.9 * TBATS_for$mean + 0.1 * NN_for$mean

combo_ts <- ts(
  combo_mean,
  start = start(test_load_ts),
  frequency = frequency(test_load_ts)
)


```

## 3.4 TBATS + NN Combination (60/40 weighting)

```{r}
#averaging of functions (60/40 split)
combo2_mean <- 0.6 * TBATS_for$mean + 0.4 * NN_for$mean

combo2_mean <- ts(
  combo2_mean,
  start = start(test_load_ts),
  frequency = frequency(test_load_ts)
)
```


## SARIMAX (Humidity Only)

```{r}
# Merge humidity
train_data <- left_join(train_data, humid_proc, by = "date")
test_data  <- left_join(test_data, humid_proc, by = "date")

xreg_train <- as.matrix(train_data %>% select(-date, -daily_ave))
xreg_test  <- as.matrix(test_data %>% select(-date, -daily_ave))

sarimax_hum_fit <- auto.arima(
  train_data$daily_ave,
  xreg = xreg_train,
  seasonal = TRUE
)

sarimax_hum_for <- forecast(
  sarimax_hum_fit,
  xreg = xreg_test,
  h = 31
)

sarimax_hum_ts <- ts(
  sarimax_hum_for$mean,
  start = end(train_load_ts) + c(0,1),
  frequency = frequency(train_load_ts)
)
```

# 4. Model Comparison
```{r}
combo2 <- accuracy(combo2_mean, test_load_ts)
TBATS_scores <- accuracy(TBATS_for$mean, test_load_ts)
combo3 <- accuracy(combo3_mean, test_load_ts)
combo <- accuracy(combo_ts, test_load_ts)
sarimax_scores <- accuracy(sarimax_hum_ts, test_load_ts)

scores <- as.data.frame(
  rbind(combo2, TBATS_scores, combo3, combo, sarimax_scores)
)

row.names(scores) <- c("TBATS+NN (60/40)","TBATS","TBATS+NN (50/50)","TBATS+NN (90/10)","SARIMAX Humidity")

kbl(scores, caption = "Forecast Accuracy for Load") %>%
  kable_styling(full_width = FALSE)
```

# 5. Forecast Comparison Plot
```{r}
autoplot(test_load_ts) +
  autolayer(TBATS_for$mean, series="TBATS") +
  autolayer(combo2_mean, series="TBATS+NN (60/40)") +
  autolayer(combo_ts, series="TBATS+NN (90/10)") +
  autolayer(sarimax_hum_ts, series="SARIMAX Humidity") +
  autolayer(combo3_mean, series = "TBATS+NN (50/50)")
  ylab("Daily Load") +
  ggtitle("Forecast Comparison (Test Period)")
```

# 6. Results and Discussion
The results indicate that seasonal structure is the dominant driver of electrical load. In the Kaggle competition the 50/50 averaged weighting for the TBATS model outperformed all other models, showing that combinations of models would improve forecasting. 

While the humidity based SARIMAX model was a top performer, incorporating weather variables through SARIMAX did not improve performance significantly compared with the TBATS+NN combination models, suggesting that weather effects are either already embedded in seasonal patterns or not effectively captured through the PCA based approach.

# 7. Git-Hub Link

https://github.com/shens2746-tech/Gutkowski_Shen_ENV797_TSA_ForecastCompetition_S26.git


# 8. AI Statement
AI tools were used to assist with debugging code and providing ideas for model implementation. All modeling decisions and interpretations were checked by group members.



