

## Gutkowski Shen Forecast Submission

Framework: 
- model + forecast
- ^repeat for all models tested / only have the code chunks showing for the models that were top 5
- comparison of all the error metrics for each model

## Cleaning Data

```{r, results = 'hide', message = FALSE}

## loading libraries

library(forecast)
library(lubridate)
library(ggplot2)
library(forecast)
library(Kendall)
library(tseries)
library(tidyverse)
library(readxl)
library(dplyr)


```

```{r}
#Inspecting Raw Data

load_raw <- read_xlsx("./Data/Raw/load.xlsx")
humid_raw <- read_xlsx("./Data/Raw/relative_humidity.xlsx")
temp_raw <- read_xlsx("./Data/Raw/temperature.xlsx")

head(load_raw)
head(humid_raw)
head(temp_raw)


```

Load data has hourly observations organized by columns, while the temperature and humidity data have hourly data in columns. For load data will merge across columns, but for the temp + humidity will group by date and average the group values by column. 


```{r}
## Cleaning Load Data

load_proc <- load_raw %>% 
  mutate(daily_ave = rowMeans(load_raw[ ,c(3:26)], na.rm = TRUE)) %>% 
  select(date, daily_ave)


head(load_proc)
tail(load_proc)


```

Dataset is cleaned between Jan 2005 - June 2011. For the project, we need to forecast January and Februaru 2011. To do this, we will need to create a time series object for January 2005 until December 2010.

```{r}
# Cleaning Humidity Data

humid_proc <- humid_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))
    

tail(humid_proc)

```


```{r}

# Cleaning Temperature Data

temp_proc <- temp_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))

#removing NA row at the end of the data
temp_rows <- nrow(temp_proc) - 1
temp_proc <- temp_proc[1:temp_rows, ]

temp_proc
tail(temp_proc)

```


## Fitting Models



# Creating Time Series Objects - Multiple Seasonality

```{r}
load_data <- load_proc

full_load_ts <- msts(load_data$daily_ave, 
                     seasonal.periods = c(7, 365.25), 
                      start=c(2005,01,01))

# creating a testing subset
n_for <- 31
full_len <- nrow(load_data)
train_data <- load_data[1:(full_len - n_for), ]
test_data  <- load_data[(full_len - n_for + 1):full_len, ]

train_load_ts <- msts(train_data$daily_ave,
                      seasonal.periods = c(7, 365.25),
                      start = c(2005, 1, 1))

test_load_ts <- msts(test_data$daily_ave,
                    seasonal.periods = c(7, 365.25),
                    start = end(train_load_ts) + c(0,1))
```


## testing models



## STL + ETS Model

```{r}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(train_load_ts,h=31)

#Plot foresting results
autoplot(ETS_fit) + ylab("Load")

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Load")
```
 
## ARIMA + Fourier terms model

```{r}

ARIMA_Four_fit <- auto.arima(train_load_ts, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(train_load_ts, 
                                          K=c(2,12))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(train_load_ts,
                                        K=c(2,12),
                                        h=31),
                           h=31
                           ) 

#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Load")

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Load")


```

 
## TBATS Model

```{r}

# TBATS can take time to fit
TBATS_fit <- tbats(train_load_ts)

TBATS_for <- forecast(TBATS_fit, h=31)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Load") 

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)+
  ylab("Load") 


```



## Neural Network Fit
```{r}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(train_load_ts,
                 p=1,
                 P=0,
                 xreg=fourier(train_load_ts, K=c(2,12)))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_fit, 
                   h=31,
                   xreg=fourier(train_load_ts, 
                                          K=c(2,12),h=31))


#Plot foresting results
autoplot(NN_for) +
  ylab("Load") 


#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Load") 
```





# Keep This Section separate / add in other models above!
# keep adding new model fits to the score tables
## will add a table that summarizes only the ones that we have submitted


## comparing forecasts
```{r}
#Model 1: STL + ETS
ETS_scores <- accuracy(ETS_fit$mean,test_load_ts)  

#Model 2: ARIMA + Fourier 
ARIMA_scores <- accuracy(ARIMA_Four_for$mean,test_load_ts)

# Model 3:  TBATS 
TBATS_scores <- accuracy(TBATS_for$mean,test_load_ts)

# Model 4:  Neural Network 
NN_scores <- accuracy(NN_for$mean,test_load_ts)


```


#comparing performance metrics

```{r}
#create data frame
scores <- as.data.frame(
  rbind(ETS_scores, ARIMA_scores, TBATS_scores, NN_scores)
  )
row.names(scores) <- c("STL+ETS", "ARIMA+Fourier","TBATS","NN")

#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))     
```

## creating a table of scores
```{r}
kbl(scores, 
      caption = "Forecast Accuracy for Load",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(scores[,"RMSE"]))
```


#plotting the different functions
```{r}
autoplot(train_load_ts) +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for,PI=FALSE, series="TBATS") +
  autolayer(NN_for,PI=FALSE, series="NN") +
  xlab("Day") + ylab("Daily Active Power") +
  guides(colour=guide_legend(title="Forecast"))
```

## taking a closer look at the forecasts
```{r}
autoplot(test_load_ts) +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for,PI=FALSE, series="TBATS") +
  autolayer(NN_for,PI=FALSE, series="NN")+
  ylab("Daily Active Power") +
  guides(colour=guide_legend(title="Forecast"))
``` 
Out of the four multi-seasonality forecast models, the TBATS model had the best performance across metrics except for MPE. For the first forecast submission, the TBATS model will be selected. 



