---
title: "ENV 797 Forecasting Competition"
subtitle: "Team 4"
author: "Gutkowski Shen"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---
## Gutkowski Shen Forecast Submission

Framework: 
- model + forecast
- ^repeat for all models tested / only have the code chunks showing for the models that were top 5
- comparison of all the error metrics for each model

## Cleaning Data

```{r, results = 'hide', message = FALSE}

## loading libraries

library(forecast)
library(lubridate)
library(ggplot2)
library(forecast)
library(Kendall)
library(tseries)
library(tidyverse)
library(readxl)
library(dplyr)

```

```{r}
#Inspecting Raw Data
load_raw <- read_xlsx("./Data/Raw/load.xlsx")
humid_raw <- read_xlsx("./Data/Raw/relative_humidity.xlsx")
temp_raw <- read_xlsx("./Data/Raw/temperature.xlsx")

head(load_raw)
head(humid_raw)
head(temp_raw)
```

Load data has hourly observations organized by columns, while the temperature and humidity data have hourly data in columns. For load data will merge across columns, but for the temp + humidity will group by date and average the group values by column. 


```{r}
## Cleaning Load Data

load_proc <- load_raw %>% 
  mutate(daily_ave = rowMeans(load_raw[ ,c(3:26)], na.rm = TRUE)) %>% 
  select(date, daily_ave)


head(load_proc)
tail(load_proc)


```

Dataset is cleaned between Jan 2005 - June 2011. For the project, we need to forecast January and Februaru 2011. To do this, we will need to create a time series object for January 2005 until December 2010.

```{r}
# Cleaning Humidity Data

humid_proc <- humid_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))
    

tail(humid_proc)

```


```{r}

# Cleaning Temperature Data

temp_proc <- temp_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))

#removing NA row at the end of the data
temp_rows <- nrow(temp_proc) - 1
temp_proc <- temp_proc[1:temp_rows, ]

temp_proc
tail(temp_proc)

```


## Fitting Models



# Creating Time Series Objects - Multiple Seasonality

```{r}
load_data <- load_proc

full_load_ts <- msts(load_data$daily_ave, 
                     seasonal.periods = c(7, 365.25), 
                      start=c(2005,01,01))

# creating a testing subset
n_for <- 31
full_len <- nrow(load_data)
train_data <- load_data[1:(full_len - n_for), ]
test_data  <- load_data[(full_len - n_for + 1):full_len, ]

train_load_ts <- msts(train_data$daily_ave,
                      seasonal.periods = c(7, 365.25),
                      start = c(2005, 1, 1))

test_load_ts <- msts(test_data$daily_ave,
                    seasonal.periods = c(7, 365.25),
                    start = end(train_load_ts) + c(0,1))
```


## testing models

## ARIMA + Fourier + Temperature Model
```{r}
# daily mean temperature
temp_daily <- temp_proc %>% 
  mutate(daily_temp = rowMeans(across(-date), na.rm = TRUE)) %>% 
  select(date, daily_temp)

# Align temperature with load training
train_temp <- temp_daily$daily_temp[1:(full_len - n_for)]
test_temp  <- temp_daily$daily_temp[(full_len - n_for + 1):full_len]

# Fourier terms + temperature
train_xreg <- cbind(fourier(train_load_ts, K = c(2, 12)), temp = train_temp)
test_xreg  <- cbind(fourier(train_load_ts, K = c(2, 12), h = 31), temp = test_temp)

ARIMA_Temp_fit <- auto.arima(train_load_ts,
                             seasonal = FALSE,
                             lambda = 0,
                             xreg = train_xreg)

ARIMA_Temp_for <- forecast(ARIMA_Temp_fit,
                           xreg = test_xreg,
                           h = 31)

#Plot forecasting results
autoplot(ARIMA_Temp_for) + ylab("Load")

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(ARIMA_Temp_for, series = "ARIMA + Fourier + Temp", PI = FALSE) +
  ylab("Load")
```

## Neural Network + Fourier + Temperature

```{r}
set.seed(797)
NN_T_fit <- nnetar(train_load_ts, p = 1, P = 0, xreg = train_xreg)

NN_T_for <- forecast(NN_T_fit, h = n_for, xreg = test_xreg)
NN_T_scores <- accuracy(NN_T_for$mean, test_load_ts)

autoplot(test_load_ts, series = "Observed") +
  autolayer(NN_T_for, series = "NN+Four+Temp", PI = FALSE) +
  ylab("Load")
```


# Keep This Section separate / add in other models above!
# keep adding new model fits to the score tables
## will add a table that summarizes only the ones that we have submitted


## comparing forecasts
```{r}

# Model 5: ARIMA + Fourier + Temperature
ARIMA_Temp_scores <- accuracy(ARIMA_Temp_for$mean, test_load_ts)

# Model 6: Neural Network + Fourier + Temperature
NN_T_scores <- accuracy(NN_T_for$mean, test_load_ts)

```


#comparing performance metrics

```{r}
#create data frame

scores <- as.data.frame(
  rbind(ARIMA_Temp_scores, NN_T_scores)
)
row.names(scores) <- c("ARIMA+Fourier+Temp", "NN+Fourier+Temp")

#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))     
```

## creating a table of scores
```{r}
kbl(scores, 
      caption = "Forecast Accuracy for Load",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(scores[,"RMSE"]))
```


## Retrain on Full Data

```{r}

final_load_ts <- full_load_ts
full_temp <- temp_daily$daily_temp[1:length(final_load_ts)]

h_fin <- 31

july_climatology <- temp_daily %>%
  filter(month(date) == 7) %>%
  mutate(md = format(date, "%m-%d")) %>%
  group_by(md) %>%
  summarise(temp_mean = mean(daily_temp, na.rm = TRUE)) %>%
  arrange(md) %>%
  pull(temp_mean)

final_xreg_tr  <- cbind(fourier(final_load_ts, K = c(2, 12)), temp = full_temp)
final_xreg_fut <- cbind(fourier(final_load_ts, K = c(2, 12), h = h_fin),
                        temp = july_climatology)
```

## Final ARIMA + Fourier + Temp

```{r}
final_ARIMA_T_fit <- auto.arima(final_load_ts, seasonal = FALSE, lambda = 0,
                                xreg = final_xreg_tr)
final_ARIMA_T <- forecast(final_ARIMA_T_fit, xreg = final_xreg_fut, h = h_fin)
autoplot(final_ARIMA_T) + ylab("Daily Load")
```

## Fianl NN + Fourier + Temp

```{r}
set.seed(797)
final_NN_T_fit <- nnetar(final_load_ts, p = 1, P = 0, xreg = final_xreg_tr)
final_NN_T <- forecast(final_NN_T_fit, h = h_fin, xreg = final_xreg_fut)
autoplot(final_NN_T) + ylab("Daily Load")
```


```{r}
dir.create("./Forecast_Submissions", showWarnings = FALSE)
template <- read_csv("submission_template.csv")

write_csv(template %>% mutate(load = as.numeric(final_ARIMA_T$mean)),
          "./Forecast_Submissions/ARIMA_Temp_submission.csv")

write_csv(template %>% mutate(load = as.numeric(final_NN_T$mean)),
          "./Forecast_Submissions/NN_Temp_submission.csv")

```


