

## Gutkowski Shen Forecast Submission

Framework: 
- model + forecast
- ^repeat for all models tested / only have the code chunks showing for the models that were top 5
- comparison of all the error metrics for each model

## Cleaning Data

```{r, results = 'hide', message = FALSE}

## loading libraries

library(forecast)
library(lubridate)
library(ggplot2)
library(forecast)
library(Kendall)
library(tseries)
library(tidyverse)
library(readxl)
library(dplyr)


```

```{r}
#Inspecting Raw Data

load_raw <- read_xlsx("./Data/Raw/load.xlsx")
humid_raw <- read_xlsx("./Data/Raw/relative_humidity.xlsx")
temp_raw <- read_xlsx("./Data/Raw/temperature.xlsx")

head(load_raw)
head(humid_raw)
head(temp_raw)


```

Load data has hourly observations organized by columns, while the temperature and humidity data have hourly data in columns. For load data will merge across columns, but for the temp + humidity will group by date and average the group values by column. 


```{r}
## Cleaning Load Data

load_proc <- load_raw %>% 
  mutate(daily_ave = rowMeans(load_raw[ ,c(3:26)], na.rm = TRUE)) %>% 
  select(date, daily_ave)


head(load_proc)
tail(load_proc)


```

Dataset is cleaned between Jan 2005 - June 2011. For the project, we need to forecast January and Februaru 2011. To do this, we will need to create a time series object for January 2005 until December 2010.

```{r}
# Cleaning Humidity Data

humid_proc <- humid_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))
    

tail(humid_proc)

```


```{r}

# Cleaning Temperature Data

temp_proc <- temp_raw %>% 
  group_by(date) %>% 
  summarise(across(3:26, ~ mean(.x, na.rm = TRUE)))

#removing NA row at the end of the data
temp_rows <- nrow(temp_proc) - 1
temp_proc <- temp_proc[1:temp_rows, ]

temp_proc
tail(temp_proc)

```


## Fitting Models



# Creating Time Series Objects - Multiple Seasonality

```{r}
load_data <- load_proc

full_load_ts <- msts(load_data$daily_ave, 
                     seasonal.periods = c(7, 365.25), 
                      start=c(2005,01,01))

# creating a testing subset
n_for <- 31
full_len <- nrow(load_data)
train_data <- load_data[1:(full_len - n_for), ]
test_data  <- load_data[(full_len - n_for + 1):full_len, ]

train_load_ts <- msts(train_data$daily_ave,
                      seasonal.periods = c(7, 365.25),
                      start = c(2005, 1, 1))

test_load_ts <- msts(test_data$daily_ave,
                    seasonal.periods = c(7, 365.25),
                    start = end(train_load_ts) + c(0,1))
```


## testing models



## STL + ETS Model

```{r}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(train_load_ts,h=31)

#Plot foresting results
autoplot(ETS_fit) + ylab("Load")

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Load")
```
 
## ARIMA + Fourier terms model

```{r}

ARIMA_Four_fit <- auto.arima(train_load_ts, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(train_load_ts, 
                                          K=c(2,12))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(train_load_ts,
                                        K=c(2,12),
                                        h=31),
                           h=31
                           ) 

#Plot foresting results
autoplot(ARIMA_Four_for) + ylab("Load")

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Load")


```

 
## TBATS Model

```{r}

# TBATS can take time to fit
TBATS_fit <- tbats(train_load_ts)

TBATS_for <- forecast(TBATS_fit, h=31)

#Plot foresting results
autoplot(TBATS_for) +
  ylab("Load") 

#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(TBATS_for, series="TBATS",PI=FALSE)+
  ylab("Load") 


```



## Neural Network Fit
```{r}
#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(train_load_ts,
                 p=1,
                 P=0,
                 xreg=fourier(train_load_ts, K=c(2,12)))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_fit, 
                   h=31,
                   xreg=fourier(train_load_ts, 
                                          K=c(2,12),h=31))


#Plot foresting results
autoplot(NN_for) +
  ylab("Load") 


#Plot model + observed data
autoplot(train_load_ts) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Load") 
```


## Incorporating Weather data
Took the first principal component from the temperature and humidity data to attempt to capture the dominant patterns in the weather variables for use as exogenous predictors in the SARIMAX model.

```{r}
temp_pc1 <- prcomp(temp_data %>% select(-date), scale. = TRUE)$x[,1]
hum_pc1  <- prcomp(humid_data %>% select(-date), scale. = TRUE)$x[,1]

weather_df <- data.frame(
  date = temp_data$date,
  temp = temp_pc1,
  humidity = hum_pc1
)

train_data <- left_join(train_data, weather_df, by = "date")
test_data  <- left_join(test_data, weather_df, by = "date")


xreg_train <- as.matrix(train_data %>% select(temp, humidity))
xreg_test  <- as.matrix(test_data %>% select(temp, humidity))

sarimax_fit <- auto.arima(
  train_data$daily_ave,
  xreg = xreg_train,
  seasonal = TRUE,
  stepwise = FALSE,
  approximation = FALSE
)

sarimax_for <- forecast(
  sarimax_fit,
  xreg = xreg_test,
  h = 31
)

sarimax_ts <- ts(
  sarimax_for$mean,
  start = end(train_load_ts) + c(0, 1),
  frequency = frequency(train_load_ts)
)

```


# Just temperature SARIMAX
Just used the first principal component of the temperature data as an exogenous predictor variable. 

```{r}
sarimax_temp_fit <- auto.arima(
  train_data$daily_ave,
  xreg = train_data$temp,
  seasonal = TRUE,
  stepwise = FALSE,
  approximation = FALSE
)

sarimax_temp_for <- forecast(
  sarimax_temp_fit,
  xreg = test_data$temp,
  h = 31
)

sarimax_temp_ts <- ts(
  sarimax_temp_for$mean,
  start = end(train_load_ts) + c(0,1),
  frequency = frequency(train_load_ts)
)



```



#just humidity sarimax
Just used the first principal component of the humidity data as an exogenous predictor variable. 

```{r}

sarimax_hum_fit <- auto.arima(
  train_data$daily_ave,
  xreg = train_data$humidity,
  seasonal = TRUE,
  stepwise = FALSE,
  approximation = FALSE
)

sarimax_hum_for <- forecast(
  sarimax_hum_fit,
  xreg = test_data$humidity,
  h = 31
)

sarimax_hum_ts <- ts(
  sarimax_hum_for$mean,
  start = end(train_load_ts) + c(0,1),
  frequency = frequency(train_load_ts)
)



```



## Combining TBATS and NN

```{r}
#averaging of functions
ensemble_mean <- 0.9 * TBATS_for$mean + 0.1 * NN_for$mean

ensemble_ts <- ts(
  ensemble_mean,
  start = start(test_load_ts),
  frequency = frequency(test_load_ts)
)
```

## Combining TBATS and NN

```{r}
#averaging of functions
combo2_mean <- 0.6 * TBATS_for$mean + 0.4 * NN_for$mean

combo2_mean <- ts(
  combo2_mean,
  start = start(test_load_ts),
  frequency = frequency(test_load_ts)
)
```

## comparing forecasts
```{r}
#Model 1: STL + ETS
ETS_scores <- accuracy(ETS_fit$mean,test_load_ts)  

#Model 2: ARIMA + Fourier 
ARIMA_scores <- accuracy(ARIMA_Four_for$mean,test_load_ts)

# Model 3:  TBATS 
TBATS_scores <- accuracy(TBATS_for$mean,test_load_ts)

# Model 4:  Neural Network 
NN_scores <- accuracy(NN_for$mean,test_load_ts)

#SARIMAX 1
sarimax_scores <- accuracy(sarimax_ts,test_load_ts)

#averaging test
tbats_nn_mix_score <- accuracy(ensemble_ts, test_load_ts)

#Just temp sarimax
temp_sarimax <- accuracy(sarimax_temp_ts, test_load_ts)

#just humidity sarimax
humid_sarimax <- accuracy(sarimax_hum_ts, test_load_ts)

#averaging2 test
tbats_nn_mix2_score <- accuracy(combo2_mean, test_load_ts)


```


#comparing performance metrics

```{r}
#create data frame
scores <- as.data.frame(
  rbind(ETS_scores, ARIMA_scores, TBATS_scores, NN_scores, sarimax_scores, tbats_nn_mix_score, temp_sarimax, humid_sarimax, tbats_nn_mix2_score)
  )
row.names(scores) <- c("STL+ETS", "ARIMA+Fourier","TBATS","NN", "Sarimax", "TBATS+NN(90/10)", "Temp SARIMAX", "Humid SARIMAX", "TBATS+NN(60/40)")

#choose model with lowest RMSE
best_model_index <- which.min(scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(scores[best_model_index,]))     
```


## creating a table of scores
```{r}
kbl(scores, 
      caption = "Forecast Accuracy for Load",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", latex_options = "hold_position") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(scores[,"RMSE"]))
```


#plotting the different functions

```{r}
autoplot(train_load_ts) +
  autolayer(test_load_ts, series = "Actual") +
  autolayer(ETS_fit, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for, PI=FALSE, series="TBATS") +
  autolayer(NN_for, PI=FALSE, series="NN") +
  autolayer(sarimax_ts, PI=FALSE, series="SARIMAX (Both)") +
  autolayer(sarimax_temp_ts, PI=FALSE, series="Temp SARIMAX") +
  autolayer(sarimax_hum_ts, PI=FALSE, series="Humid SARIMAX") +
  autolayer(ensemble_ts, PI=FALSE, series="TBATS+NN") +
  xlab("Day") + 
  ylab("Daily Active Power") +
  ggtitle("Model Comparison - Full Series") +
  guides(colour = guide_legend(title = "Forecast"))
```



## taking a closer look at the forecasts
```{r}
autoplot(test_load_ts) +
  autolayer(ETS_fit$mean, PI=FALSE, series="STL+ETS") +
  autolayer(ARIMA_Four_for$mean, PI=FALSE, series="ARIMA + Fourier") +
  autolayer(TBATS_for$mean, PI=FALSE, series="TBATS") +
  autolayer(NN_for$mean, PI=FALSE, series="NN") +
  autolayer(sarimax_ts, PI=FALSE, series="SARIMAX (Temp and Humidity)") +
  autolayer(sarimax_temp_ts, PI=FALSE, series="Temp SARIMAX") +
  autolayer(sarimax_hum_ts, PI=FALSE, series="Humid SARIMAX") +
  autolayer(ensemble_ts, PI=FALSE, series="TBATS+NN") +
  ylab("Daily Active Power") +
  ggtitle("Forecast Comparison - Test Period") +
  guides(colour = guide_legend(title = "Forecast"))
``` 





