---
title: "NG_Forecast"
output:
  pdf_document: default
  html_document: default
date: "2026-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
library(lubridate)
library(ggplot2)
library(forecast)
library(Kendall)
library(tseries)
library(tidyverse)
library(readxl)
library(dplyr)
library(cowplot)

```

Initial Steps
1. plots
2. acf/ Pacf
3. trend
4. Seasonality
5. Stationary tests
6. Outliers
7. Missing data



```{r}
#full data until July 31
load_data <- read_csv("./Data/Model_Training/Load_Train.csv")


#subset of data until June 30 (used for training the mode) - July 1-31 we will be forecasting
train_rows <- nrow(load_data) - 31


load_subset <- load_data[1:train_rows, ]

#checking the data ends on the right date
tail(load_data)
tail(load_subset)
head(load_subset)

```



# Creating Time Series Objects - Multiple Seasonality

```{r}
full_load_ts <- msts(load_data$daily_ave, 
                     seasonal.periods = c(7, 365.25), 
                      start=c(2005,01,01))

# creating a testing subset
n_for <- 31
full_len <- nrow(load_data)
train_data <- load_data[1:(full_len - n_for), ]
test_data  <- load_data[(full_len - n_for + 1):full_len, ]

train_load_ts <- msts(train_data$daily_ave,
                      seasonal.periods = c(7, 365.25),
                      start = c(2005, 1, 1))

test_load_ts <- msts(test_data$daily_ave,
                    seasonal.periods = c(7, 365.25),
                    start = end(train_load_ts) + c(0,1))
```


```{r}
#Checking to make sure that the testing/ training subsets were lined up in the correct locations. Full ts was plotted underneath layers. 

autoplot(full_load_ts, series = "full ts")+
  autolayer(train_load_ts, series = "training ts")+
  autolayer(test_load_ts, series = "test")

```

```{r}
train_load_ts_arima <- ts(train_data$daily_ave, frequency = 7)
full_load_ts_arima  <- ts(load_data$daily_ave, frequency = 7)

fit_arima <- auto.arima(train_load_ts_arima)
fc_arima <- forecast(fit_arima, h = 31)
accuracy(fc_arima, test_data$daily_ave)
```

# Explaination:
RMSE and MAE are substantially higher on the test set compared to the training set, showing that the predictive accuracy is dropped
The MAPE on the test set is aorund 13.52%, showing that on average, the forecast deviates from the true values by about 13.52%.
The MASE value exceeds 1 (≈1.67), indicating that the ARIMA model performs worse than a naive benchmark model on the test set.

```{r}
fit_final <- auto.arima(full_load_ts_arima)
fc_final <- forecast(fit_final, h = 31)
```
 
```{r}
template <- read_csv("~/TSA_SP26/group/Data/Model_Training/submission_template.csv")
head(template)
names(template)
```
```{r}
submission <- template %>%
  mutate(load = as.numeric(fc_final$mean))

write_csv(submission, "~/TSA_SP26/group/Data/Model_Training/my_submission.csv")
```

## calculating ACF and PACF (here I'm not sure should we use train_load_ts or train_load_ts_arima, but the result seems like the same)

```{r}
load_acf <- autoplot(Acf(train_load_ts, lag.max = 40, plot = FALSE))
load_pacf <- autoplot(Pacf(train_load_ts, lag.max = 40,  plot = FALSE))


plot_grid(load_acf, load_pacf, nrow = 1)
```

```{r}
load_acf <- autoplot(Acf(train_load_ts_arima, lag.max = 40, plot = FALSE))
load_pacf <- autoplot(Pacf(train_load_ts_arima, lag.max = 40, plot = FALSE))

plot_grid(load_acf, load_pacf, nrow = 1)
```

# decomposing the series to assess the seasonal/ trend components

```{r}
train_load_ts %>% mstl() %>%
  autoplot()
```
```{r}
## Stationarity test
adf.test(train_load_ts_arima)
```

```{r}
## Outlier visualization

autoplot(train_load_ts_arima) +
  ggtitle("Load Time Series")
```




